---
title: "bmi-gender"
author: "Zhongyun Zhang"
date: "2023-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## bmi

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tableone)
library(tidyverse)
library(MatchIt)
library(lmtest)
library(sandwich)
library(dplyr)
library("viridis")
library(gridExtra)
#install.packages("patchwork")
library(patchwork)
library(forestplot)
library(grid)
#install.packages("forestploter")
library(forestploter)
```

```{r}
#import data
eicu <- read.csv("~/Desktop/301/PPC/data/hf_eicu.csv",row.names = 1, header = TRUE)
mimic <- read.csv("~/Desktop/301/PPC/data/hf_mimic.csv", row.names = 1, header = TRUE)
#View(eicu)
```

```{r}
eicu <- subset(eicu,eicu$numbedscategory==">= 500" | eicu$numbedscategory=="250 - 499")
```

```{r}
## fill na for MV hours
eicu[is.na(eicu)] <- 0
mimic[is.na(mimic)] <- 0
# Remove rows with one patient with 300 in 'age' - mimic
mimic <- mimic[mimic$age != "300", ]

#select variables
vars <- c('admission_type', 'gender', 'age','bmi', 'charlson_comorbidity_index', 'ethnicity', 'cabg', 'valve_surgery', 'sofa', 'vent_hour', 'los_hospital_day', 'los_icu_day')
eicu <- eicu %>% select(all_of(vars))
mimic <- mimic %>% select(all_of(vars))
# combine eicu and mimic
total <- rbind(eicu, mimic) %>%
  filter(ethnicity %in% c('black', 'white'))
#hist(total$ethnicity,total$age)
total <- total[total$vent_hour >= 0,]
```

```{r}
total$bmi_c <-  ifelse(total$bmi < 18.5, "Underweight",
                 ifelse(total$bmi >= 18.5 & total$bmi < 25, "Healthy weight",
                 ifelse(total$bmi >= 25 & total$bmi < 30, "Overweight",
                 ifelse(total$bmi >= 30 & total$bmi < 40, "Obese",                                             "Extremely obese"))))

# Count the frequencies of each BMI category
bmi_counts <- table(total$bmi_c)
bmi_counts <- bmi_counts[c("Underweight", "Healthy weight", "Overweight", "Obese","Extremely obese")]
```

```{r}
# Plot the bar chart
plot1 <- barplot(bmi_counts, 
        xlab = "BMI category", 
        ylab = "Patient number", 
        las = 1.5, 
        cex.names = 0.7, 
        ylim = c(0,4000),
        col = "skyblue", 
        border = "black",
        cex.axis = 0.7, cex.lab = 0.7,width = 0.3) 
```

## Gender
## MV HOUR 24

```{r}
#male as 1 female as 0
#total$gender <- factor(ifelse(total$gender == "M", 1, 0))
#vent hour over 24 as 1, under 24 as 0
total$vent24 <- factor(ifelse(total$vent_hour > 24, 1, 0))
total$bmi_c <- factor(total$bmi_c, levels = c("Underweight", "Healthy weight", "Overweight", "Obese", "Extremely obese"))
levels(total$bmi_c)
total$bmi_c <- relevel(total$bmi_c, ref = "Healthy weight")
```


```{r}
total$gender <- factor(total$gender)
#levels(total$gender)
total$gender <- relevel(total$gender, ref = "M")
```

```{r}
total$ethnicity <- factor(total$ethnicity)
#levels(total$ethnicity)
total$ethnicity <- relevel(total$ethnicity, ref = "white")
```


### Underweight

```{r}
vent_24_1 <- glm(vent24 ~ gender+age+ethnicity+admission_type+charlson_comorbidity_index +cabg+valve_surgery+sofa, data =subset(total, bmi_c == "Underweight"), family = binomial)
sum_vent24_1 <- summary(vent_24_1)
sum_vent24_1
```

```{r}
invfisher.logit <- vcov(vent_24_1)
CI.logit = vent_24_1$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),t(t(sqrt(diag(invfisher.logit)))))
out.mv.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.mv.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
#rownames(out.mv.f) = c('bmi_cUnderweight','bmi_cOverweight', 'bmi_cObesity,class I','bmi_cObesity,class II', 'bmi_cObesity,class III ',  'age', 'ethnicityblack','genderM','admission_type','charlson_comorbidity_index','cabg','valve_surgery','sofa')
rownames(out.mv.f) = c( 'genderF', 'age','ethnicityblack','admission_type',
                        'charlson_comorbidity_index',
                        'cabg','valve_surgery','sofa')
out.mv.f
odd_241 <- data.frame(out.mv.f[1, ])
# Print the new dataframe
print(odd_241)
```


### Healthy Weight

```{r}
vent_24_2 <- glm(vent24 ~ gender+age+ethnicity+admission_type+charlson_comorbidity_index +cabg+valve_surgery+sofa, data =subset(total, bmi_c == "Healthy weight"), family = binomial)
sum_vent24_2 <- summary(vent_24_2)
sum_vent24_2
```

```{r}
invfisher.logit <- vcov(vent_24_2)
CI.logit = vent_24_2$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),t(t(sqrt(diag(invfisher.logit)))))
out.mv.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.mv.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
#rownames(out.mv.f) = c('bmi_cUnderweight','bmi_cOverweight', 'bmi_cObesity,class I','bmi_cObesity,class II', 'bmi_cObesity,class III ',  'age', 'ethnicityblack','genderM','admission_type','charlson_comorbidity_index','cabg','valve_surgery','sofa')
rownames(out.mv.f) = c( 'genderF','age','ethnicityblack', 'admission_type',
                        'charlson_comorbidity_index',
                        'cabg','valve_surgery','sofa')
out.mv.f
odd_242 <- data.frame(out.mv.f[1, ])
# Print the new dataframe
print(odd_242)
```

### Over Weight

```{r}
vent_24_3 <- glm(vent24 ~ gender+age+ethnicity+admission_type+charlson_comorbidity_index +cabg+valve_surgery+sofa, data =subset(total, bmi_c == "Overweight"), family = binomial)
sum_vent24_3 <- summary(vent_24_3)
sum_vent24_3
```

```{r}
invfisher.logit <- vcov(vent_24_3)
CI.logit = vent_24_3$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),t(t(sqrt(diag(invfisher.logit)))))
out.mv.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.mv.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
#rownames(out.mv.f) = c('bmi_cUnderweight','bmi_cOverweight', 'bmi_cObesity,class I','bmi_cObesity,class II', 'bmi_cObesity,class III ',  'age', 'ethnicityblack','genderM','admission_type','charlson_comorbidity_index','cabg','valve_surgery','sofa')
rownames(out.mv.f) = c( 'genderF','age','ethnicityblack', 'admission_type',
                        'charlson_comorbidity_index',
                        'cabg','valve_surgery','sofa')
out.mv.f
odd_243 <- data.frame(out.mv.f[1, ])
# Print the new dataframe
print(odd_243)
```

### Obese

```{r}
vent_24_4 <- glm(vent24 ~ gender+age+ethnicity+admission_type+charlson_comorbidity_index +cabg+valve_surgery+sofa, data =subset(total, bmi_c == "Obese"), family = binomial)
sum_vent24_4 <- summary(vent_24_4)
sum_vent24_4
```

```{r}
invfisher.logit <- vcov(vent_24_4)
CI.logit = vent_24_4$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),t(t(sqrt(diag(invfisher.logit)))))
out.mv.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.mv.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
#rownames(out.mv.f) = c('bmi_cUnderweight','bmi_cOverweight', 'bmi_cObesity,class I','bmi_cObesity,class II', 'bmi_cObesity,class III ',  'age', 'ethnicityblack','genderM','admission_type','charlson_comorbidity_index','cabg','valve_surgery','sofa')
rownames(out.mv.f) = c( 'genderF','age','ethnicityblack', 'admission_type',
                        'charlson_comorbidity_index',
                        'cabg','valve_surgery','sofa')
out.mv.f
odd_244 <- data.frame(out.mv.f[1, ])
# Print the new dataframe
print(odd_244)
```


### Extremely obese

```{r}
vent_24_5 <- glm(vent24 ~ gender+age+ethnicity+admission_type+charlson_comorbidity_index +cabg+valve_surgery+sofa, data =subset(total, bmi_c == "Extremely obese"), family = binomial)
sum_vent24_5 <- summary(vent_24_5)
sum_vent24_5
```

```{r}
invfisher.logit <- vcov(vent_24_5)
CI.logit = vent_24_5$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),t(t(sqrt(diag(invfisher.logit)))))
out.mv.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.mv.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
#rownames(out.mv.f) = c('bmi_cUnderweight','bmi_cOverweight', 'bmi_cObesity,class I','bmi_cObesity,class II', 'bmi_cObesity,class III ',  'age', 'ethnicityblack','genderM','admission_type','charlson_comorbidity_index','cabg','valve_surgery','sofa')
rownames(out.mv.f) = c( 'genderF','age','ethnicityblack', 'admission_type',
                        'charlson_comorbidity_index',
                        'cabg','valve_surgery','sofa')
out.mv.f
odd_245 <- data.frame(out.mv.f[1, ])
# Print the new dataframe
print(odd_245)
```
### Table count

```{r}
df_mv24_1 <- total %>%
  group_by(bmi_c, gender) %>%
  count()%>%
  pivot_wider(names_from = gender, values_from = n)%>%
  rename(Male = M, Female = F, "BMI" = bmi_c)

df_mv24_1
```

```{r}
df_mv24_2 <- total %>%
  group_by(bmi_c, gender, vent24) %>%
  filter(vent24 == 0)%>%
  count() %>%
  pivot_wider(names_from = gender, values_from = n) %>%
  rename(Male = M, Female = F, "BMI" = bmi_c) %>%
  select("BMI", "Male", "Female")
# Print the updated data frame
#print(df_mv24_2)
```

```{r}
combined_df1 <- inner_join(df_mv24_2, df_mv24_1, by = c("BMI"))

# Connect the corresponding numbers with "/"
combined_df1$Male <- paste(combined_df1$Male.x, combined_df1$Male.y, sep = "/")
combined_df1$Female <- paste(combined_df1$Female.x, combined_df1$Female.y, sep = "/")

# Remove the unnecessary columns
combined_df1 <- combined_df1 %>%
  select("BMI", "Male", "Female") %>%
  select(-vent24)  
# Print the combined dataframe
combined_df1 <- print(combined_df1[2:4])
#combined_df1
```

```{r}
## Combine tables
#mv24_table <- cbind(combined_df1,mv24_or,p_values_df)
desired_order <- c("Underweight", "Healthy weight", "Overweight", "Obese", "Extremely obese")

# Reorder the rows based on the desired order
combined_df1 <- combined_df1 %>%
  mutate(BMI = factor(BMI, levels = desired_order)) %>%
  arrange(BMI)
```


### Odd Combined

```{r}
mv24_odd <- t(cbind(odd_241,odd_242,odd_243,odd_244,odd_245))
new_col_names <- c("OddsRatio", "Lower_CI", "Upper_CI")  
colnames(mv24_odd) <- new_col_names 
new_row_names <- c("Underweight", "Healthy weight", "Overweight", "Obese", "Extremely obese") 
rownames(mv24_odd) <- new_row_names 
mv24_odd <- data.frame(mv24_odd)
mv24_odd
```

### P-value

```{r}
mv24_odd$"OR(95% CI)" <- sprintf(
  "%.2f (%.2f, %.2f)",
  mv24_odd$OddsRatio,
  mv24_odd$Lower_CI,
  mv24_odd$Upper_CI
)

# Extract the p-values for the "genderF" variable from each regression
p_values <- c(sum_vent24_1$coefficients["genderF", "Pr(>|z|)"],
              sum_vent24_2$coefficients["genderF", "Pr(>|z|)"],
              sum_vent24_3$coefficients["genderF", "Pr(>|z|)"],
              sum_vent24_4$coefficients["genderF", "Pr(>|z|)"],
              sum_vent24_5$coefficients["genderF", "Pr(>|z|)"])

# Create a dataframe with one column for the p-values
p_values_df <- data.frame(ifelse(p_values > 0.05, ">0.05", ifelse(p_values < 0.01, "<0.01", round(p_values, 3))))
colnames(p_values_df) <- "P value"


# Print the dataframe
#print(p_values_df)
table_mv <- cbind(combined_df1,mv24_odd,p_values_df)
table_mv$'               '<- "               "
colnames(table_mv)[ncol(table_mv)] <- "               "
print(table_mv)
```

### Forest Plot

```{r}
tm <- forest_theme(base_size = 8,           # 设置基础字体大小
                   refline_col = "red4",     # 设置参考线颜色为红色
                   arrow_type = "closed",    # 设置箭头类型为闭合箭头
                   footnote_col = "blue4",
                   colnames_just = "center")   
# 绘制森林图
p1 <- forest(table_mv[, c("BMI","Male","Female","OR(95% CI)","P value",'               ')],   # 选择要在森林图中使用的数据列，这里包括变量名列、患者数量列、绘图要用的空白列和HR（95%CI）列
            est = table_mv$OddsRatio,          # 效应值，也就是HR列
            lower = table_mv$Lower_CI,  # 置信区间下限
            upper = table_mv$Upper_CI,  # 置信区间上限
            sizes = 0.2,        # 黑框框的大小
            ci_column = 6,             # 在第3列（可信区间列）绘制森林图
            ref_line = 1,              # 添加参考线
#            arrow_lab = c("Low risk", "High Risk"),  # 箭头标签，用来表示效应方向，如何设置取决于你的样本情况
            xlim = c(0,3),          # 设置x轴范围
            ticks_at = c(0,1,2,3),  # 在指定位置添加刻度
            theme = tm)               # 添加自定义主题
#            footnote = "This is the demo data. Please feel free to change\nanything you want.")  # 添加脚注信息
p1
```

## ICU Day 3

```{r}
#icu day over 3 as 1, under 3 as 0
total$icu3 <- factor(ifelse(total$los_icu_day > 3, 1, 0))
```

### Underweight

```{r}
icu_3_1 <- glm(icu3 ~ gender+age+ethnicity+admission_type+charlson_comorbidity_index +cabg+valve_surgery+sofa, data =subset(total, bmi_c == "Underweight"), family = binomial)
sum_icu3_1 <- summary(icu_3_1)
sum_icu3_1
```

```{r}
invfisher.logit <- vcov(icu_3_1)
CI.logit = icu_3_1$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),t(t(sqrt(diag(invfisher.logit)))))
out.mv.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.mv.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
#rownames(out.mv.f) = c('bmi_cUnderweight','bmi_cOverweight', 'bmi_cObesity,class I','bmi_cObesity,class II', 'bmi_cObesity,class III ',  'age', 'ethnicityblack','genderM','admission_type','charlson_comorbidity_index','cabg','valve_surgery','sofa')
rownames(out.mv.f) = c( 'genderF', 'age','ethnicityblack','admission_type',
                        'charlson_comorbidity_index',
                        'cabg','valve_surgery','sofa')
out.mv.f
odd_31 <- data.frame(out.mv.f[1, ])
# Print the new dataframe
print(odd_31)
```


### Healthy Weight

```{r}
icu_3_2 <- glm(icu3 ~ gender+age+ethnicity+admission_type+charlson_comorbidity_index +cabg+valve_surgery+sofa, data =subset(total, bmi_c == "Healthy weight"), family = binomial)
sum_icu3_2 <- summary(icu_3_2)
sum_icu3_2
```

```{r}
invfisher.logit <- vcov(icu_3_2)
CI.logit = icu_3_2$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),t(t(sqrt(diag(invfisher.logit)))))
out.mv.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.mv.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
#rownames(out.mv.f) = c('bmi_cUnderweight','bmi_cOverweight', 'bmi_cObesity,class I','bmi_cObesity,class II', 'bmi_cObesity,class III ',  'age', 'ethnicityblack','genderM','admission_type','charlson_comorbidity_index','cabg','valve_surgery','sofa')
rownames(out.mv.f) = c( 'genderF', 'age','ethnicityblack','admission_type',
                        'charlson_comorbidity_index',
                        'cabg','valve_surgery','sofa')
out.mv.f
odd_32 <- data.frame(out.mv.f[1, ])
# Print the new dataframe
print(odd_32)
```


### Overweight

```{r}
icu_3_3 <- glm(icu3 ~ gender+age+ethnicity+admission_type+charlson_comorbidity_index +cabg+valve_surgery+sofa, data =subset(total, bmi_c == "Overweight"), family = binomial)
sum_icu3_3 <- summary(icu_3_3)
sum_icu3_3
```

```{r}
invfisher.logit <- vcov(icu_3_3)
CI.logit = icu_3_3$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),t(t(sqrt(diag(invfisher.logit)))))
out.mv.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.mv.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
#rownames(out.mv.f) = c('bmi_cUnderweight','bmi_cOverweight', 'bmi_cObesity,class I','bmi_cObesity,class II', 'bmi_cObesity,class III ',  'age', 'ethnicityblack','genderM','admission_type','charlson_comorbidity_index','cabg','valve_surgery','sofa')
rownames(out.mv.f) = c( 'genderF', 'age','ethnicityblack','admission_type',
                        'charlson_comorbidity_index',
                        'cabg','valve_surgery','sofa')
out.mv.f
odd_33 <- data.frame(out.mv.f[1, ])
# Print the new dataframe
print(odd_33)
```

### Obese

```{r}
icu_3_4 <- glm(icu3 ~ gender+age+ethnicity+admission_type+charlson_comorbidity_index +cabg+valve_surgery+sofa, data =subset(total, bmi_c == "Obese"), family = binomial)
sum_icu3_4 <- summary(icu_3_4)
sum_icu3_4
```

```{r}
invfisher.logit <- vcov(icu_3_4)
CI.logit = icu_3_4$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),t(t(sqrt(diag(invfisher.logit)))))
out.mv.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.mv.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
#rownames(out.mv.f) = c('bmi_cUnderweight','bmi_cOverweight', 'bmi_cObesity,class I','bmi_cObesity,class II', 'bmi_cObesity,class III ',  'age', 'ethnicityblack','genderM','admission_type','charlson_comorbidity_index','cabg','valve_surgery','sofa')
rownames(out.mv.f) = c( 'genderF', 'age','ethnicityblack','admission_type',
                        'charlson_comorbidity_index',
                        'cabg','valve_surgery','sofa')
out.mv.f
odd_34 <- data.frame(out.mv.f[1, ])
# Print the new dataframe
print(odd_34)
```

### Extremely Obese

```{r}
icu_3_5 <- glm(icu3 ~ gender+age+ethnicity+admission_type+charlson_comorbidity_index +cabg+valve_surgery+sofa, data =subset(total, bmi_c == "Extremely obese"), family = binomial)
sum_icu3_5 <- summary(icu_3_5)
sum_icu3_5
```

```{r}
invfisher.logit <- vcov(icu_3_5)
CI.logit = icu_3_5$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),t(t(sqrt(diag(invfisher.logit)))))
out.mv.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.mv.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
#rownames(out.mv.f) = c('bmi_cUnderweight','bmi_cOverweight', 'bmi_cObesity,class I','bmi_cObesity,class II', 'bmi_cObesity,class III ',  'age', 'ethnicityblack','genderM','admission_type','charlson_comorbidity_index','cabg','valve_surgery','sofa')
rownames(out.mv.f) = c( 'genderF', 'age','ethnicityblack','admission_type',
                        'charlson_comorbidity_index',
                        'cabg','valve_surgery','sofa')
out.mv.f
odd_35 <- data.frame(out.mv.f[1, ])
# Print the new dataframe
print(odd_35)
```

### Table count

```{r}
df_icu3_1 <- total %>%
  group_by(bmi_c, gender) %>%
  count()%>%
  pivot_wider(names_from = gender, values_from = n)%>%
  rename(Male = M, Female = F, "BMI" = bmi_c)

df_icu3_1
```

```{r}
df_icu3_2 <- total %>%
  group_by(bmi_c, gender, icu3) %>%
  filter(icu3 == 0)%>%
  count() %>%
  pivot_wider(names_from = gender, values_from = n) %>%
  rename(Male = M, Female = F, "BMI" = bmi_c) %>%
  select("BMI", "Male", "Female")
# Print the updated data frame
print(df_icu3_2)
```

```{r}
combined_df2 <- inner_join(df_icu3_2, df_icu3_1, by = c("BMI"))

# Connect the corresponding numbers with "/"
combined_df2$Male <- paste(combined_df2$Male.x, combined_df2$Male.y, sep = "/")
combined_df2$Female <- paste(combined_df2$Female.x, combined_df2$Female.y, sep = "/")

# Remove the unnecessary columns
combined_df2 <- combined_df2 %>%
  select("BMI", "Male", "Female") %>%
  select(-icu3)  
# Print the combined dataframe
combined_df2 <- print(combined_df2[2:4])
combined_df2
```

```{r}
## Combine tables
#icu3_table <- cbind(combined_df2,icu3_or)
desired_order <- c("Underweight", "Healthy weight", "Overweight", "Obese", "Extremely obese")

# Reorder the rows based on the desired order
combined_df2 <- combined_df2 %>%
  mutate(BMI = factor(BMI, levels = desired_order)) %>%
  arrange(BMI)
```

### Odd Combined

```{r}
icu3_odd <- t(cbind(odd_31,odd_32,odd_33,odd_34,odd_35))
new_col_names <- c("OddsRatio", "Lower_CI", "Upper_CI")  
colnames(icu3_odd) <- new_col_names 
new_row_names <- c("Underweight", "Healthy weight", "Overweight", "Obese", "Extremely obese")  
rownames(icu3_odd) <- new_row_names 
icu3_odd <- data.frame(icu3_odd)
icu3_odd
```

### P-value

```{r}
icu3_odd$"OR(95% CI)" <- sprintf(
  "%.2f (%.2f, %.2f)",
  icu3_odd$OddsRatio,
  icu3_odd$Lower_CI,
  icu3_odd$Upper_CI
)
icu3_table <- cbind(combined_df2,icu3_odd,p_values_df)
print(icu3_table)
```

```{r}
# Extract the p-values for the "genderF" variable from each regression
p_values <- c(sum_icu3_1$coefficients["genderF", "Pr(>|z|)"],
              sum_icu3_2$coefficients["genderF", "Pr(>|z|)"],
              sum_icu3_3$coefficients["genderF", "Pr(>|z|)"],
              sum_icu3_4$coefficients["genderF", "Pr(>|z|)"],
              sum_icu3_5$coefficients["genderF", "Pr(>|z|)"])

# Create a dataframe with one column for the p-values
p_values_df <- data.frame(ifelse(p_values > 0.05, ">0.05", ifelse(p_values < 0.01, "<0.01", round(p_values, 3))))
colnames(p_values_df) <- "P value"

# Print the dataframe
#print(p_values_df)

# Print the dataframe
table_icu <- cbind(combined_df2,icu3_odd,p_values_df)
table_icu$'               '<- "               "
colnames(table_icu)[ncol(table_icu)] <- "               "
print(table_icu)
```

### forest plot

```{r}
tm <- forest_theme(base_size = 8,           # 设置基础字体大小
                   refline_col = "red4",     # 设置参考线颜色为红色
                   arrow_type = "closed",    # 设置箭头类型为闭合箭头
                   footnote_col = "blue4",
                   colnames_just = "center")   
# 绘制森林图
p2 <- forest(table_icu[, c("BMI","Male","Female","OR(95% CI)","P value",'               ')],   # 选择要在森林图中使用的数据列，这里包括变量名列、患者数量列、绘图要用的空白列和HR（95%CI）列
            est = table_icu$OddsRatio,          # 效应值，也就是HR列
            lower = table_icu$Lower_CI,  # 置信区间下限
            upper = table_icu$Upper_CI,  # 置信区间上限
            sizes = 0.2,        # 黑框框的大小
            ci_column = 6,             # 在第3列（可信区间列）绘制森林图
            ref_line = 1,              # 添加参考线
#            arrow_lab = c("Low risk", "High Risk"),  # 箭头标签，用来表示效应方向，如何设置取决于你的样本情况
            xlim = c(0,3),          # 设置x轴范围
            ticks_at = c(0,1,2,3),  # 在指定位置添加刻度
            theme = tm)               # 添加自定义主题
#            footnote = "This is the demo data. Please feel free to change\nanything you want.")  # 添加脚注信息
p2

```


## Hospital Day 7

```{r}
#hos day over 7 as 1, under 7 as 0
total$hos7 <- factor(ifelse(total$los_hospital_day > 7, 1, 0))
```

### Underweight

```{r}
hos_7_1 <- glm(hos7 ~ gender+age+ethnicity+admission_type+charlson_comorbidity_index +cabg+valve_surgery+sofa, data =subset(total, bmi_c == "Underweight"), family = binomial)
sum_hos7_1 <- summary(hos_7_1)
sum_hos7_1
```

```{r}
invfisher.logit <- vcov(hos_7_1)
CI.logit = hos_7_1$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),t(t(sqrt(diag(invfisher.logit)))))
out.mv.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.mv.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
#rownames(out.mv.f) = c('bmi_cUnderweight','bmi_cOverweight', 'bmi_cObesity,class I','bmi_cObesity,class II', 'bmi_cObesity,class III ',  'age', 'ethnicityblack','genderM','admission_type','charlson_comorbidity_index','cabg','valve_surgery','sofa')
rownames(out.mv.f) = c( 'genderF', 'age','ethnicityblack','admission_type',
                        'charlson_comorbidity_index',
                        'cabg','valve_surgery','sofa')
out.mv.f
odd_71 <- data.frame(out.mv.f[1, ])
# Print the new dataframe
print(odd_71)
```


### Healthy Weight

```{r}
hos_7_2 <- glm(hos7 ~ gender+age+ethnicity+admission_type+charlson_comorbidity_index +cabg+valve_surgery+sofa, data =subset(total, bmi_c == "Healthy weight"), family = binomial)
sum_hos7_2 <- summary(hos_7_2)
sum_hos7_2
```

```{r}
invfisher.logit <- vcov(hos_7_2)
CI.logit = hos_7_2$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),t(t(sqrt(diag(invfisher.logit)))))
out.mv.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.mv.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
#rownames(out.mv.f) = c('bmi_cUnderweight','bmi_cOverweight', 'bmi_cObesity,class I','bmi_cObesity,class II', 'bmi_cObesity,class III ',  'age', 'ethnicityblack','genderM','admission_type','charlson_comorbidity_index','cabg','valve_surgery','sofa')
rownames(out.mv.f) = c( 'genderF', 'age','ethnicityblack','admission_type',
                        'charlson_comorbidity_index',
                        'cabg','valve_surgery','sofa')
out.mv.f
odd_72 <- data.frame(out.mv.f[1, ])
# Print the new dataframe
print(odd_72)
```

### Overweight

```{r}
hos_7_3 <- glm(hos7 ~ gender+age+ethnicity+admission_type+charlson_comorbidity_index +cabg+valve_surgery+sofa, data =subset(total, bmi_c == "Overweight"), family = binomial)
sum_hos7_3 <- summary(hos_7_3)
sum_hos7_3
```

```{r}
invfisher.logit <- vcov(hos_7_3)
CI.logit = hos_7_3$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),t(t(sqrt(diag(invfisher.logit)))))
out.mv.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.mv.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
#rownames(out.mv.f) = c('bmi_cUnderweight','bmi_cOverweight', 'bmi_cObesity,class I','bmi_cObesity,class II', 'bmi_cObesity,class III ',  'age', 'ethnicityblack','genderM','admission_type','charlson_comorbidity_index','cabg','valve_surgery','sofa')
rownames(out.mv.f) = c( 'genderF', 'age','ethnicityblack','admission_type',
                        'charlson_comorbidity_index',
                        'cabg','valve_surgery','sofa')
out.mv.f
odd_73 <- data.frame(out.mv.f[1, ])
# Print the new dataframe
print(odd_73)
```


### Obese

```{r}
hos_7_4 <- glm(hos7 ~ gender+age+ethnicity+admission_type+charlson_comorbidity_index +cabg+valve_surgery+sofa, data =subset(total, bmi_c == "Obese"), family = binomial)
sum_hos7_4 <- summary(hos_7_4)
sum_hos7_4
```

```{r}
invfisher.logit <- vcov(hos_7_4)
CI.logit = hos_7_4$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),t(t(sqrt(diag(invfisher.logit)))))
out.mv.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.mv.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
#rownames(out.mv.f) = c('bmi_cUnderweight','bmi_cOverweight', 'bmi_cObesity,class I','bmi_cObesity,class II', 'bmi_cObesity,class III ',  'age', 'ethnicityblack','genderM','admission_type','charlson_comorbidity_index','cabg','valve_surgery','sofa')
rownames(out.mv.f) = c( 'genderF', 'age','ethnicityblack','admission_type',
                        'charlson_comorbidity_index',
                        'cabg','valve_surgery','sofa')
out.mv.f
odd_74 <- data.frame(out.mv.f[1, ])
# Print the new dataframe
print(odd_74)
```

### Extremely Obese

```{r}
hos_7_5 <- glm(hos7 ~ gender+age+ethnicity+admission_type+charlson_comorbidity_index +cabg+valve_surgery+sofa, data =subset(total, bmi_c == "Extremely obese"), family = binomial)
sum_hos7_5 <- summary(hos_7_5)
sum_hos7_5
```

```{r}
invfisher.logit <- vcov(hos_7_5)
CI.logit = hos_7_5$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),t(t(sqrt(diag(invfisher.logit)))))
out.mv.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.mv.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
#rownames(out.mv.f) = c('bmi_cUnderweight','bmi_cOverweight', 'bmi_cObesity,class I','bmi_cObesity,class II', 'bmi_cObesity,class III ',  'age', 'ethnicityblack','genderM','admission_type','charlson_comorbidity_index','cabg','valve_surgery','sofa')
rownames(out.mv.f) = c( 'genderF', 'age','ethnicityblack','admission_type',
                        'charlson_comorbidity_index',
                        'cabg','valve_surgery','sofa')
out.mv.f
odd_75 <- data.frame(out.mv.f[1, ])
# Print the new dataframe
print(odd_75)
```

### Table count

```{r}
df_hos7_1 <- total %>%
  group_by(bmi_c, gender) %>%
  count()%>%
  pivot_wider(names_from = gender, values_from = n)%>%
  rename(Male = M, Female = F, "BMI" = bmi_c)

df_hos7_1
```

```{r}
df_hos7_2 <- total %>%
  group_by(bmi_c, gender, hos7) %>%
  filter(hos7 == 0)%>%
  count() %>%
  pivot_wider(names_from = gender, values_from = n) %>%
  rename(Male = M, Female = F, "BMI" = bmi_c) %>%
  select("BMI", "Male", "Female")
# Print the updated data frame
print(df_hos7_2)
```

```{r}
combined_df3 <- inner_join(df_hos7_2, df_hos7_1, by = c("BMI"))

# Connect the corresponding numbers with "/"
combined_df3$Male <- paste(combined_df3$Male.x, combined_df3$Male.y, sep = "/")
combined_df3$Female <- paste(combined_df3$Female.x, combined_df3$Female.y, sep = "/")

# Remove the unnecessary columns
combined_df3 <- combined_df3 %>%
  select("BMI", "Male", "Female") %>%
  select(-hos7)  
# Print the combined dataframe
combined_df3 <- print(combined_df3[2:4])
combined_df3
```

```{r}
## Combine tables
#hos7_table <- cbind(combined_df3,)
desired_order <- c("Underweight", "Healthy weight", "Overweight", "Obese", "Extremely obese")

# Reorder the rows based on the desired order
combined_df3 <- combined_df3 %>%
  mutate(BMI = factor(BMI, levels = desired_order)) %>%
  arrange(BMI)
```

### Odd Combined

```{r}
hos7_odd <- t(cbind(odd_71,odd_72,odd_73,odd_74,odd_75))
new_col_names <- c("OddsRatio", "Lower_CI", "Upper_CI")  
colnames(hos7_odd) <- new_col_names 
new_row_names <- c("Underweight", "Healthy weight", "Overweight", "Obese", "Extremely obese")
rownames(hos7_odd) <- new_row_names 
hos7_odd <- data.frame(hos7_odd)
hos7_odd
```

### P-value

```{r}
hos7_odd$"OR(95% CI)" <- sprintf(
  "%.2f (%.2f, %.2f)",
  hos7_odd$OddsRatio,
  hos7_odd$Lower_CI,
  hos7_odd$Upper_CI
)
```


```{r}
# Extract the p-values for the "genderF" variable from each regression
p_values <- c(sum_hos7_1$coefficients["genderF", "Pr(>|z|)"],
              sum_hos7_2$coefficients["genderF", "Pr(>|z|)"],
              sum_hos7_3$coefficients["genderF", "Pr(>|z|)"],
              sum_hos7_4$coefficients["genderF", "Pr(>|z|)"],
              sum_hos7_5$coefficients["genderF", "Pr(>|z|)"])

# Create a dataframe with one column for the p-values
p_values_df <- data.frame(ifelse(p_values > 0.05, ">0.05", ifelse(p_values < 0.01, "<0.01", round(p_values, 3))))
colnames(p_values_df) <- "P value"

table_hos <- cbind(combined_df3,hos7_odd, p_values_df)
table_hos$'               '<- "               "
colnames(table_hos)[ncol(table_hos)] <- "               "
print(table_hos)
print(table_hos)
```

### forest plot

```{r}
tm <- forest_theme(base_size = 8,           # 设置基础字体大小
                   refline_col = "red4",     # 设置参考线颜色为红色
                   arrow_type = "closed",    # 设置箭头类型为闭合箭头
                   footnote_col = "blue4",
                   colnames_just_each = "center")   
# 绘制森林图
p3 <- forest(table_hos[, c("BMI","Male","Female","OR(95% CI)","P value",'               ')],   # 选择要在森林图中使用的数据列，这里包括变量名列、患者数量列、绘图要用的空白列和HR（95%CI）列
            est = table_hos$OddsRatio,          # 效应值，也就是HR列
            lower = table_hos$Lower_CI,  # 置信区间下限
            upper = table_hos$Upper_CI,  # 置信区间上限
            sizes = 0.2,        # 黑框框的大小
            ci_column = 6,             # 在第3列（可信区间列）绘制森林图
            ref_line = 1,              # 添加参考线
#            arrow_lab = c("Low risk", "High Risk"),  # 箭头标签，用来表示效应方向，如何设置取决于你的样本情况
            xlim = c(0,3),          # 设置x轴范围
            ticks_at = c(0,1,2,3),  # 在指定位置添加刻度
            theme = tm)               # 添加自定义主题
#            footnote = "This is the demo data. Please feel free to change\nanything you want.")  # 添加脚注信息
p3

```
## Save Table

```{r}
ggsave("mv24_bmi_gender.png", plot = p1)
ggsave("icu3_bmi_gender.png", plot = p2)
ggsave("hos7_bmi_gender.png", plot = p3)
```


