---
title: "Inequality Study - PSM"
output: pdf_document
date: '2023-05-10'
author: 'Shihui Zhu'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tableone)
library(tidyverse)
library(MatchIt)
library(lmtest)
library(sandwich)
library(dplyr)
```

## Read in data

* Outcomes: vent hours, hospital days, icu days
* Controlling variables: age, bmi, CCI score, SOFA score, Surgery type (cabg, valve)
* Varaibles: race (black and white), gender

```{r}
#import data
eicu <- read.csv("~/Desktop/301/PPC/data/hf_eicu.csv",row.names = 1, header = TRUE)
mimic <- read.csv("~/Desktop/301/PPC/data/hf_mimic.csv", row.names = 1, header = TRUE)
```

## Select eICU patients beds number > 250, MIMIC age outlier

```{r}
eicu <- eicu %>% dplyr::filter(numbedscategory %in% c(">= 500", "250 - 499"))
mimic <- mimic %>% dplyr::filter(age != 300)
```

## Select variables

```{r}
## fill na for MV hours
eicu[is.na(eicu)] <- 0
mimic[is.na(mimic)] <- 0

vars <- c('admission_type','bmi', 'gender', 'age', 'charlson_comorbidity_index', 'ethnicity', 'cabg', 'valve_surgery', 'sofa', 'vent_hour', 'los_hospital_day', 'los_icu_day')

eicu <- eicu %>% select(all_of(vars))
mimic <- mimic %>% select(all_of(vars))

total <- rbind(eicu, mimic) %>%
  filter(ethnicity %in% c('black', 'white'))
  # vent hour -> mv > 24, mv <= 24
```

total$mv <- factor(ifelse(total$vent_hour > 24, 1, 0))
total$los_hospital_day <- factor(ifelse(total$los_hospital_day > 7, 1, 0))
total$los_icu_day <- factor(ifelse(total$los_icu_day > 3, 1, 0))
#total <- total %>% select(-vent_hour)


```{r}
total$bmi_c <-  ifelse(total$bmi < 18.5, "Underweight",
                 ifelse(total$bmi >= 18.5 & total$bmi < 25, "Healthy Weight",
                 ifelse(total$bmi >= 25 & total$bmi < 30, "Overweight",
                 ifelse(total$bmi >= 30 & total$bmi < 40, "Obese",                                             "Extremely Obese"))))


total$mv_group <-  ifelse(total$vent_hour == 0.000, "0",
                 ifelse(total$vent_hour > 0 & total$vent_hour <= 12, "12",
                 ifelse(total$vent_hour > 12 & total$vent_hour <= 24, "24",
                 ifelse(total$vent_hour > 24 & total$vent_hour <= 48,"48",
                 ifelse(total$vent_hour > 48 & total$vent_hour <= 168, "168",">168")))))

total$los_icu_group <-  ifelse(total$los_icu_day <= 2, "2",
                 ifelse(total$los_icu_day > 2 & total$los_icu_day <= 3, "3",
                 ifelse(total$los_icu_day > 3 & total$los_icu_day <= 7, "7",
                 ifelse(total$los_icu_day > 7 & total$los_icu_day <= 14, "14",">14"))))

total$los_hospital_group <-  ifelse(total$los_hospital_day <= 3, "3",
                 ifelse(total$los_hospital_day > 3 & total$los_hospital_day <= 7, "7",
                 ifelse(total$los_hospital_day > 7 & total$los_hospital_day <= 14, "14",
                 ifelse(total$los_hospital_day > 14 & total$los_hospital_day <= 30, "30", 
                        ">30"))))
```

```{r}
# Reorder levels of age_group variable
#total$age_group <- factor(total$age_group, levels = c("18-45", "45-65", "65-75",">75"))

# Reorder levels of mv_group variable
total$mv_group <- factor(total$mv_group, levels = c("0", "12", "24", "48","168",">168"))

# Reorder levels of los_icu_group variable
total$los_icu_group <- factor(total$los_icu_group, levels = c("2", "3", "7", "14", ">14"))

# Reorder levels of los_hospital_group variable
total$los_hospital_group <- factor(total$los_hospital_group, levels = c("3", "7", "14", "30", ">30"))

total$bmi_c <- factor(total$bmi_c, levels = c("Underweight", "Healthy Weight", "Overweight", "Obese", "Extremely Obese"))
```

## TableOne - before PSM with SMD

```{r}
# by gender
table2 <- CreateTableOne(vars = c('age','ethnicity', 'bmi_c','admission_type',  'charlson_comorbidity_index', 'sofa', 'cabg', 'valve_surgery', 'mv_group', 'los_icu_group', 'los_hospital_group'),
                         data = total,
                         factorVars = c('admission_type',
                                        'ethnicity',"bmi_c", 'cabg', 
                                        'valve_surgery', 'mv_group', 
                                        'los_icu_group',
                                        'los_hospital_group'),
                         strata = 'gender',
                         smd = TRUE)

table2 <- print(table2, smd=TRUE, showAllLevels = TRUE, noSpaces = TRUE, 
                printToggle = FALSE, nonnormal = c("charlson_comorbidity_index", 'sofa','age'))
write.csv(table2, file="~/Desktop/301/PPC/output/table1_bygender_beforematching.csv")
```

```{r}
# by race
table3 <- CreateTableOne(vars = c('age','gender', 'bmi_c','admission_type',  'charlson_comorbidity_index', 'sofa', 'cabg', 'valve_surgery', 'mv_group', 'los_icu_group', 'los_hospital_group'),
                         data = total,
                         factorVars = c('admission_type', 'bmi_c',
                                        'gender', 'cabg', 
                                        'valve_surgery', 
                                        'mv_group', 
                                        'los_icu_group',
                                        'los_hospital_group'),
                         strata = 'ethnicity',
                         smd = TRUE)

table3 <- print(table3, smd=TRUE, showAllLevels = TRUE, noSpaces = TRUE, 
                printToggle = FALSE, nonnormal = c("charlson_comorbidity_index", 'sofa','age'))
write.csv(table3, file="~/Desktop/301/PPC/output/table1_byrace_beforematching.csv")
```


## PSM

0. Prepare 

```{r}
total[,c('admission_type', 'cabg', 'valve_surgery', 'gender', 'ethnicity')] <- 
  lapply(total[,c('admission_type', 'cabg', 'valve_surgery', 'gender', 'ethnicity')], factor)
total$ethnicity <- relevel(total$ethnicity, ref='white')
total$gender <- relevel(total$gender, ref='M')
```



1. PSM matching


1. Gender (control for ...)

```{r}
match_obj <- matchit(gender ~ age+ethnicity+bmi+admission_type+
                         charlson_comorbidity_index+sofa+cabg+valve_surgery, data = total, 
                     method = "nearest", distance = "glm",
                     ratio = 1,
                     replace = FALSE)
summary(match_obj, standardize = T)$sum.matched
```

MAX SMD < 0.1

### Plot the match result

```{r}
plot(match_obj, type = "jitter", interactive = FALSE, cex.main = 1)
plot(summary(match_obj), abs = FALSE)
```

2. Race (control for ...)

```{r}
match_obj_1 <- matchit(ethnicity ~ age+gender+bmi+admission_type+
                         charlson_comorbidity_index+sofa+cabg+valve_surgery, data = total,
                       method = "nearest", distance = "glm",
                    ratio = 1,
                    replace = FALSE)
summary(match_obj_1, standardize = T)$sum.matched
```

* MAX SMD < 0.1

### Plot the match result

```{r}
plot(match_obj_1, type = "jitter", interactive = FALSE)
plot(summary(match_obj_1), abs = FALSE)
```

## Extract Matched data

```{r}
matched_data <- match.data(match_obj)
head(matched_data)

matched_data_1 <- match.data(match_obj_1)
head(matched_data_1)
```

## baseline covariates of matched data

* test to see if varaibles are balanced

```{r}
# by gender
table4 <- CreateTableOne (vars = c('age','ethnicity', 'bmi_c','admission_type',  'charlson_comorbidity_index', 'sofa', 'cabg', 'valve_surgery', 'mv_group', 'los_icu_group', 'los_hospital_group'),
                          data = matched_data,
                          factorVars = c('admission_type', "bmi_c",
                                        'cabg', 
                                        'ethnicity',
                                        'valve_surgery',
                                        'mv_group', 
                                        'los_icu_group',
                                        'los_hospital_group'),
                          strata = 'gender',
                          smd=TRUE)
table4 <- print(table4, smd=TRUE, showAllLevels = TRUE, noSpaces = TRUE, 
                printToggle = FALSE, nonnormal = c("charlson_comorbidity_index", 'sofa','age'))
write.csv(table4, 
          file="~/Desktop/301/PPC/output/table1_bygender_aftermatching.csv")
```

```{r}
# by race
table5 <- CreateTableOne (vars = c('age','gender', 'bmi_c','admission_type',  'charlson_comorbidity_index', 'sofa', 'cabg', 'valve_surgery', 'mv_group', 'los_icu_group', 'los_hospital_group'),
                          data = matched_data_1,
                          factorVars = c('admission_type', "bmi_c",
                                        'cabg', 
                                        'gender',
                                        'valve_surgery',
                                        'mv_group', 
                                        'los_icu_group',
                                        'los_hospital_group'),
                          strata = 'ethnicity',
                          smd=TRUE)
table5 <- print(table5, smd=TRUE, showAllLevels = TRUE, noSpaces = TRUE, 
                printToggle = FALSE, nonnormal = c("charlson_comorbidity_index", 'sofa','age'))
write.csv(table5, file="~/Desktop/301/PPC/output/table1_byrace_aftermatching.csv")
```

### Inspection: 

1. by gender: covariates are balanced (p.value > 0.05, SMD < 0.1), all three outcomes are not the same between groups

2. by race: covariates are balanced (p.value > 0.05, SMD < 0.1), MV hours and hospital days are not the same between groups

## ATT (Average Treatment Effect on the Treated)

### LM

1. Gender - affecting hospital days, icu days, and MV hours

#### MV hours

```{r}
res <- glm(mv ~ gender, data = matched_data, 
          weights = matched_data$weights, family = binomial(link = 'logit'))
invfisher.logit <- vcov(res)
CI.logit = res$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),
                                              t(t(sqrt(diag(invfisher.logit)))))
out.mv.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.mv.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
rownames(out.mv.f) = c('female')
out.mv.f
```

95% CI is (0.1041158, 0.3800868)

#### ICU days

```{r}
res <- glm(los_icu_day ~ gender, data = matched_data, 
          weights = matched_data$weights, family = binomial(link = 'logit'))
invfisher.logit <- vcov(res)
CI.logit = res$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),
                                              t(t(sqrt(diag(invfisher.logit)))))
out.icu.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.icu.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
rownames(out.icu.f) = c('female')
out.icu.f
```

#### Hospital days

```{r}
res <- glm(los_hospital_day ~ gender, data = matched_data, 
          weights = matched_data$weights, family = binomial(link = 'logit'))
invfisher.logit <- vcov(res)
CI.logit = res$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),
                                              t(t(sqrt(diag(invfisher.logit)))))
out.hos.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.hos.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
rownames(out.hos.f) = c('female')
out.hos.f
```

2. Race - affecting hospital days

#### MV

```{r}
res <- glm(mv ~ ethnicity, data = matched_data_1, 
          weights = matched_data_1$weights, family = binomial(link = 'logit'))
invfisher.logit <- vcov(res)
CI.logit = res$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),
                                              t(t(sqrt(diag(invfisher.logit)))))
out.mv.race = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.mv.race) = c('Estimate for OR','95% CI Lower','95% CI Upper')
rownames(out.mv.race) = c('black')
out.mv.race
```

#### Hospital

```{r}
res <- glm(los_hospital_day ~ ethnicity, data = matched_data_1, 
          weights = matched_data_1$weights, family = binomial(link = 'logit'))
invfisher.logit <- vcov(res)
CI.logit = res$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),
                                              t(t(sqrt(diag(invfisher.logit)))))
out.hos.race = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.hos.race) = c('Estimate for OR','95% CI Lower','95% CI Upper')
rownames(out.hos.race) = c('black')
out.hos.race
```

#### ICU

```{r}
res <- glm(los_icu_day ~ ethnicity, data = matched_data_1, 
          weights = matched_data_1$weights, family = binomial(link = 'logit'))
invfisher.logit <- vcov(res)
CI.logit = res$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),
                                              t(t(sqrt(diag(invfisher.logit)))))
out.icu.race = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.icu.race) = c('Estimate for OR','95% CI Lower','95% CI Upper')
rownames(out.icu.race) = c('black')
out.icu.race
```

No significant differece for icu days among race group

## Summary

```{r}
out.f <- rbind(out.mv.f, out.icu.f, out.hos.f)
row.names(out.f) <- c("MV 24 hours", "ICU 3 days", "Hospital 7 days")
out.r <- rbind(out.mv.race, out.icu.race, out.hos.race)
row.names(out.r) <- c("MV 24 hours", "ICU 3 days", "Hospital 7 days")
out.f %>% knitr::kable(digits = 3)
out.r %>% knitr::kable(digits = 3)
```

All are different except for ICU days among black and whites

### Further

1. Sensitivity test?

2. Category -> 4?

## Conclusion

GLM regression of post-surgical MV hours, hospital hours, icu hours as a function of race and sex, controlling for age, CCI score, SOFA, along with procedure (cabg or valve), all included as covariates in the model.

** Vent hours -> Qualitative, PMV yes or no
** model for eICU
** model for MIMIC
** model MIMIC and eICU > 500
** by Age
** multiple ICU admissions among female patients with long MV hours

** Plots
0, 0-24, 24-48, 48-7d, >7d for MV hours
0-3, 3-7, >7d






