---
title: "inequality_4"
author: "Zhongyun Zhang"
date: "2023-06-24"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tableone)
library(tidyverse)
library(MatchIt)
library(lmtest)
library(sandwich)
library(dplyr)
library("viridis")
#install.packages("gtools", repos="https://cran.rstudio.com/")
library(gtools)
install.packages("RColorBrewer")
library(RColorBrewer)
library(gridExtra)

```

## Data Processing

* Outcomes: vent hours, hospital days, icu days
* Controlling variables: age, CCI score, SOFA score, Surgery type (cabg, valve)
* Varaibles: race (black and white), gender


```{r}
#import data
eicu <- read.csv("~/Desktop/301/PPC/data/hf_eicu.csv",row.names = 1, header = TRUE)
mimic <- read.csv("~/Desktop/301/PPC/data/hf_mimic.csv", row.names = 1, header = TRUE)
#View(eicu)
```

```{r}
eicu_500 <- subset(eicu, eicu$numbedscategory==">= 500")
eicu_250499 <- subset(eicu, eicu$numbedscategory=="250 - 499")
eicu_100249 <- subset(eicu, eicu$numbedscategory=="100 - 249")
#table(eicu_500$ethnicity)
#table(eicu_250499$ethnicity)
#table(eicu_100249$ethnicity)
#selecting eicu number of beds>250 
eicu <- subset(eicu,eicu$numbedscategory==">= 500" | eicu$numbedscategory=="250 - 499")
#table(eicu$ethnicity)
#table(mimic$ethnicity)
```

```{r}
## fill na for MV hours
eicu[is.na(eicu)] <- 0
mimic[is.na(mimic)] <- 0
# Remove rows with one patient with 300 in 'age' - mimic
mimic <- mimic[mimic$age != "300", ]

#select variables
vars <- c('admission_type', 'gender', 'age', 'charlson_comorbidity_index', 'ethnicity', 'cabg', 'valve_surgery', 'sofa', 'vent_hour', 'los_hospital_day', 'los_icu_day','bmi')
eicu <- eicu %>% select(all_of(vars))
mimic <- mimic %>% select(all_of(vars))
# combine eicu and mimic
total <- rbind(eicu, mimic) %>%
  filter(ethnicity %in% c('black', 'white'))
#hist(total$ethnicity,total$age)
total <- total[total$vent_hour >= 0,]
table(total$ethnicity)
```

```{r}
# Assuming your dataset is named 'data' and 'gender' and 'ethnicity' are existing columns
total$combined <- ifelse(total$gender == "M" & total$ethnicity == "white", "White male",
                 ifelse(total$gender == "F" & total$ethnicity == "white", "White female",
                 ifelse(total$gender == "M" & total$ethnicity == "black", "Black male",
                 ifelse(total$gender == "F" & total$ethnicity == "black", "Black female",NA))))
# Convert 'combined' column to numeric representation
total$combined <- factor(total$combined, levels = c("White male", "White female", "Black male", "Black female"), labels = c(1, 2, 3, 4))
levels(total$combined) <- c("White male", "White female", "Black male", "Black female")
table(total$combined)


#total$age_c <-  ifelse(total$age < 40, "<40",
#                 ifelse(total$age >= 40 & total$age < 50, "40-50",
#                 ifelse(total$age >= 50 & total$age < 60, "50-60",
#                 ifelse(total$age >= 60 & total$age < 70, "60-70",               
#                 ifelse(total$age >= 70 & total$age < 80, "70-80",
#                 ifelse(total$age >= 80 & total$age < 90, "80-90",">=90"))))))

#total$sofa <- as.factor(total$sofa)
#total$charlson_comorbidity_index <- as.factor(total$charlson_comorbidity_index)
#total$sofa_c <- factor(ifelse(total$sofa > 6, 1, 0))
#total$charlson_comorbidity_index_c <- factor(ifelse(total$charlson_comorbidity_index > 4, 1, 0))

                 
  # vent hour -> mv > 24, mv <= 24
total$mv24 <- factor(ifelse(total$vent_hour > 24, 1, 0))
total$hos_day3 <- factor(ifelse(total$los_hospital_day > 3, 1, 0))
total$icu_day3 <- factor(ifelse(total$los_icu_day > 3, 1, 0))
```


#### MV hours 24

```{r}
res1 <- glm(mv24 ~ combined+admission_type+age+bmi+charlson_comorbidity_index +
                    cabg+valve_surgery+sofa, data = total, family = binomial)
sum_res1 <- summary(res1)
#sum_res1
```

```{r}
invfisher.logit <- vcov(res1)
CI.logit = res1$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))), t(t(sqrt(diag(invfisher.logit)))))
out.hos.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.hos.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
rownames(out.hos.f) = c("White female", "Black male", "Black female",'admission_type','age','bmi','charlson_comorbidity_index','cabg','valve_surgery','sofa')
out.hos.f 
```

the probability for ventilation over 24 hour for white female, black male, black female would be 1.29, 2.05,2.38 times higher compared to white male respectively, with adjusted confounders age, admission type, charlson comorbidity index, cabg, valve surgery and sofa score.


```{r}
# Get the predicted probabilities
predicted <- predict(res1, newdata = total, type = "response")
lower_ci <- predict(res1, type = "response", se.fit = TRUE)$fit - 1.96 * predict(res1, type = "response", se.fit = TRUE)$se.fit 
upper_ci <- predict(res1, type = "response", se.fit = TRUE)$fit + 1.96 * predict(res1, type = "response", se.fit = TRUE)$se.fit
#results for mv24
mv24_table <- aggregate(cbind(predicted, lower_ci, upper_ci) ~ combined, data = total, FUN = mean)
# Print the aggregated data
#p_values
p_values <- sum_res1$coefficients[2:4, "Pr(>|z|)"]
p_values <- data.frame(p_values = c(NA,p_values))
rownames(p_values) <- c("White male", "White female", "Black male", "Black female")
mv24_table <- cbind(mv24_table,p_values)
print(mv24_table)
```

## mv hours 0

```{r}
total$mv0 <- factor(ifelse(total$vent_hour > 0, 1, 0))
```

```{r}
res12 <- glm(mv0 ~ combined+admission_type+age+bmi+charlson_comorbidity_index +
                    cabg+valve_surgery+sofa, data = total, family = binomial)
sum_res12 <- summary(res12)
#sum_res12
```

```{r}
invfisher.logit <- vcov(res12)
CI.logit = res12$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),
                                              t(t(sqrt(diag(invfisher.logit)))))
out.hos.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.hos.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
rownames(out.hos.f) = c("White female", "Black male", "Black female",'admission_type','age','bmi','charlson_comorbidity_index','cabg','valve_surgery','sofa')
out.hos.f
```

```{r}
# Get the predicted probabilities
predicted <- predict(res12, newdata = total, type = "response")
lower_ci <- predict(res12, type = "response", se.fit = TRUE)$fit - 1.96 * predict(res12, type = "response", se.fit = TRUE)$se.fit 
upper_ci <- predict(res12, type = "response", se.fit = TRUE)$fit + 1.96 * predict(res12, type = "response", se.fit = TRUE)$se.fit

mv0_table <- aggregate(cbind(predicted, lower_ci, upper_ci) ~ combined, data = total, FUN = mean)
# Print the aggregated data
p_values <- sum_res12$coefficients[2:4, "Pr(>|z|)"]
p_values <- data.frame(p_values = c(NA,p_values))
rownames(p_values) <- c("White male", "White female", "Black male", "Black female")
mv0_table <- cbind(mv0_table,p_values)
print(mv0_table)
```

## mv hours 12

```{r}
total$mv12 <- factor(ifelse(total$vent_hour > 12, 1, 0))
```

```{r}
res2 <- glm(mv12 ~ combined+admission_type+age+bmi+charlson_comorbidity_index +
                    cabg+valve_surgery+sofa, data = total, family = binomial)
sum_res2 <- summary(res2)
#sum_res2
```

```{r}
invfisher.logit <- vcov(res2)
CI.logit = res2$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),
                                              t(t(sqrt(diag(invfisher.logit)))))
out.hos.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.hos.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
rownames(out.hos.f) = c("White female", "Black male", "Black female",'admission_type','age','bmi','charlson_comorbidity_index','cabg','valve_surgery','sofa')
out.hos.f
```

```{r}
# Get the predicted probabilities
predicted <- predict(res2, newdata = total, type = "response")
lower_ci <- predict(res2, type = "response", se.fit = TRUE)$fit - 1.96 * predict(res2, type = "response", se.fit = TRUE)$se.fit 
upper_ci <- predict(res2, type = "response", se.fit = TRUE)$fit + 1.96 * predict(res2, type = "response", se.fit = TRUE)$se.fit

mv12_table <- aggregate(cbind(predicted, lower_ci, upper_ci) ~ combined, data = total, FUN = mean)
# Print the aggregated data
p_values <- sum_res2$coefficients[2:4, "Pr(>|z|)"]
p_values <- data.frame(p_values = c(NA,p_values))
rownames(p_values) <- c("White male", "White female", "Black male", "Black female")
mv12_table <- cbind(mv12_table,p_values)
print(mv12_table)
```


## mv hours 48

```{r}
total$mv48 <- factor(ifelse(total$vent_hour > 48, 1, 0))
```

```{r}
res3 <- glm(mv48 ~ combined+admission_type+age+bmi+charlson_comorbidity_index +
                    cabg+valve_surgery+sofa, data = total, family = binomial)
sum_res3 <- summary(res3)
#sum_res3
```

```{r}
invfisher.logit <- vcov(res3)
CI.logit = res3$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),
                                              t(t(sqrt(diag(invfisher.logit)))))
out.hos.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.hos.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
rownames(out.hos.f) = c("White female", "Black male", "Black female",'admission_type','age','bmi','charlson_comorbidity_index','cabg','valve_surgery','sofa')
out.hos.f
```

```{r}
# Get the predicted probabilities
predicted <- predict(res3, newdata = total, type = "response")
lower_ci <- predict(res3, type = "response", se.fit = TRUE)$fit - 1.96 * predict(res3, type = "response", se.fit = TRUE)$se.fit 
upper_ci <- predict(res3, type = "response", se.fit = TRUE)$fit + 1.96 * predict(res3, type = "response", se.fit = TRUE)$se.fit

mv48_table <- aggregate(cbind(predicted, lower_ci, upper_ci) ~ combined, data = total, FUN = mean)
# Print the aggregated data
p_values <- sum_res3$coefficients[2:4, "Pr(>|z|)"]
p_values <- data.frame(p_values = c(NA,p_values))
rownames(p_values) <- c("White male", "White female", "Black male", "Black female")
mv48_table <- cbind(mv48_table,p_values)
print(mv48_table)
```

## mv hours >7 DAYS (168 HOURS)

```{r}
total$mv168 <- factor(ifelse(total$vent_hour > 168, 1, 0))
```

```{r}
res13 <- glm(mv168 ~ combined+admission_type+age+bmi+charlson_comorbidity_index + cabg+valve_surgery+sofa, data = total, family = binomial)
sum_res13 <- summary(res13)
#sum_res3
```

```{r}
invfisher.logit <- vcov(res13)
CI.logit = res3$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),
                                              t(t(sqrt(diag(invfisher.logit)))))
out.hos.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.hos.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
rownames(out.hos.f) = c("White female", "Black male", "Black female",'admission_type','age','bmi','charlson_comorbidity_index','cabg','valve_surgery','sofa')
out.hos.f
```

```{r}
# Get the predicted probabilities
predicted <- predict(res13, newdata = total, type = "response")
lower_ci <- predict(res13, type = "response", se.fit = TRUE)$fit - 1.96 * predict(res13, type = "response", se.fit = TRUE)$se.fit 
upper_ci <- predict(res13, type = "response", se.fit = TRUE)$fit + 1.96 * predict(res13, type = "response", se.fit = TRUE)$se.fit

mv168_table <- aggregate(cbind(predicted, lower_ci, upper_ci) ~ combined, data = total, FUN = mean)
# Print the aggregated data
p_values <- sum_res13$coefficients[2:4, "Pr(>|z|)"]
p_values <- data.frame(p_values = c(NA,p_values))
rownames(p_values) <- c("White male", "White female", "Black male", "Black female")
mv168_table <- cbind(mv168_table,p_values)
print(mv168_table)
```

```{r}
mv0_table$mv <- "0"
mv12_table$mv <- "12"
mv24_table$mv <- "24"
mv48_table$mv <- "48"
mv168_table$mv <- "168"

combined_data <- bind_rows(mv0_table, mv12_table,mv24_table,mv48_table,mv168_table)

combined_data$result_predicted <- (1-combined_data$predicted)*100
combined_data$result_lower_ci <- (1-combined_data$upper_ci)*100
combined_data$result_upper_ci <- (1-combined_data$lower_ci)*100

                         
p_values_filtered <- combined_data[!is.na(combined_data$p_values), ]

p_values_filtered <- p_values_filtered %>% 
  mutate(p_values = round(p_values, digits = 8))
# Function to add stars to p-values less than a threshold

spacing <- 0.02
#my_palette <- c('#C1CCAE','#A9A7D6','#8DC0B1','#F2C190')
#my_palette <- c('#C1CCAE','#F8D695','#A9A7D6','#EDB3B2')
my_palette <- c('#162C9a',"#9F2218",'#81b6e3','#F07928')

```


combined_data$p_value_category <- ifelse(combined_data$p_value < 0.001, "***",
                                         ifelse(combined_data$p_value < 0.01, "**",
                                                ifelse(combined_data$p_value < 0.05, "*", "")))


## p-value stars


line_mv <- ggplot(combined_data, aes(x = factor(mv, level = c('0', '12', '24', '48','168')), y = result_predicted, group = combined, color = combined)) +
  geom_errorbar(aes(ymin = result_lower_ci, ymax = result_upper_ci), width = 0.1) +
  geom_line(linewidth = 1, aes(label = p_value_category)) +
  geom_text(aes(label = p_value_category), hjust = 0, vjust = 4, size = 4) +
  labs(x = "MV duration (hour)", y = "MV removal rate (%)") +
  scale_color_manual(name = NULL, values = my_palette) +
  scale_y_continuous(limits = c(0, 110), expand = c(0, 0)) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 12),
    legend.position = "bottom",
    panel.background = element_blank(),
    panel.grid.major = element_blank(),    # Remove major vertical grid lines
    panel.grid.minor = element_blank(),    # Remove minor vertical grid lines
    panel.border = element_blank(),
    axis.line = element_line(colour = "black", linewidth = 0.5),
    panel.grid.major.y = element_line(colour = "grey", linetype = "solid", linewidth = 0.2)) +
  guides(color = guide_legend(override.aes = list(label = "", shape = 0, linetype = 1)))  # Remove the label and keep only the color lines

line_mv


```{r}
line_mv <- ggplot(combined_data, aes(x = factor(mv, level = c('0', '12', '24', '48','168')), y = result_predicted, group = combined, color = combined)) +
  geom_errorbar(aes(ymin = result_lower_ci, ymax = result_upper_ci), width = 0.1) +
  geom_line(linewidth = 1) + 
  labs(x = "MV duration (hour)", y = "MV removal rate (%)") +
  scale_color_manual(name = NULL,values = my_palette) +
  scale_y_continuous(limits = c(0, 110), expand = c(0, 0)) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 12),
#    legend.position = "bottom",
#    legend.justification = c(1, 0),
#    legend.box.just = "right",
    panel.background = element_blank(),
    panel.grid.major = element_blank(),    # Remove major vertical grid lines
    panel.grid.minor = element_blank(),    # Remove minor vertical grid lines
    panel.border = element_blank(),
    axis.line = element_line(colour = "black", linewidth = 0.5),
    panel.grid.major.y = element_line(colour = "grey", linetype = "solid", linewidth = 0.2)  # Customize horizontal grid lines
  ) + theme(legend.position =  c(0.8, 0.25)) +
  theme(legend.background = element_rect(fill = "white", colour = "grey")) +
  theme(legend.key.height= unit(0.4, 'cm'),
        legend.key.width= unit(0.2, 'cm'))
line_mv
```


As we can see from the graph, at 0 hour which means patients remove ventilation as soon as they get into ICU, black female has the lowest probability to remove ventilation at 0 hour while white male has higher probability to remove ventilation. At 12 hour when they get into ICU, 70% of white male could remove ventilation within 12 hours, while black female has 57%, black male has 59%, and white female has 65%. For 24 hour, white male has the probability of removing ventilation within 24 hour is 89%, while white female, black male, black female has probability removing of 86%, 80%, 77% respectively.


#### icu days 3

```{r}
res4 <- glm(icu_day3 ~ combined + admission_type+age+bmi+charlson_comorbidity_index +
                    cabg+valve_surgery+sofa, data = total, family = binomial(link = 'logit'))
sum_res4 <- summary(res4)
#sum_res4
```

```{r}
invfisher.logit <- vcov(res4)
CI.logit = res4$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),
                                              t(t(sqrt(diag(invfisher.logit)))))
out.hos.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.hos.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
rownames(out.hos.f) = c("White female", "Black male", "Black female",'admission_type','age','bmi','charlson_comorbidity_index','cabg','valve_surgery','sofa')
out.hos.f
```

```{r}
# Get the predicted probabilities
predicted <- predict(res4, newdata = total, type = "response")
lower_ci <- predict(res4, type = "response", se.fit = TRUE)$fit - 1.96 * predict(res4, type = "response", se.fit = TRUE)$se.fit 
upper_ci <- predict(res4, type = "response", se.fit = TRUE)$fit + 1.96 * predict(res4, type = "response", se.fit = TRUE)$se.fit
#results for mv24
icu3_table <- aggregate(cbind(predicted, lower_ci, upper_ci) ~ combined, data = total, FUN = mean)
# Print the aggregated data
p_values <- sum_res4$coefficients[2:4, "Pr(>|z|)"]
p_values <- data.frame(p_values = c(NA,p_values))
rownames(p_values) <- c("White male", "White female", "Black male", "Black female")
icu3_table <- cbind(icu3_table,p_values)
print(icu3_table)
```

## icu days 2

```{r}
total$icu_day2 <- factor(ifelse(total$los_icu_day > 2, 1, 0))
```

```{r}
res5 <- glm(icu_day2 ~ combined + admission_type+age+bmi+charlson_comorbidity_index +
                    cabg+valve_surgery+sofa, data = total, family = binomial)
sum_res5 <- summary(res5)
#sum_res5
```

```{r}
invfisher.logit <- vcov(res5)
CI.logit = res5$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),
                                              t(t(sqrt(diag(invfisher.logit)))))
out.hos.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.hos.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
rownames(out.hos.f) = c("White female", "Black male", "Black female",'admission_type','age','bmi','charlson_comorbidity_index','cabg','valve_surgery','sofa')
out.hos.f
```

```{r}
# Get the predicted probabilities
predicted <- predict(res5, newdata = total, type = "response")
lower_ci <- predict(res5, type = "response", se.fit = TRUE)$fit - 1.96 * predict(res5, type = "response", se.fit = TRUE)$se.fit 
upper_ci <- predict(res5, type = "response", se.fit = TRUE)$fit + 1.96 * predict(res5, type = "response", se.fit = TRUE)$se.fit
#results for icu2
icu2_table <- aggregate(cbind(predicted, lower_ci, upper_ci) ~ combined, data = total, FUN = mean)
# Print the aggregated data
p_values <- sum_res5$coefficients[2:4, "Pr(>|z|)"]
p_values <- data.frame(p_values = c(NA,p_values))
rownames(p_values) <- c("White male", "White female", "Black male", "Black female")
icu2_table <- cbind(icu2_table,p_values)
print(icu2_table)
```

## icu days 7

```{r}
total$icu_day7 <- factor(ifelse(total$los_icu_day > 7, 1, 0))
```

```{r}
res6 <- glm(icu_day7 ~ combined + admission_type+age+bmi+charlson_comorbidity_index +
                    cabg+valve_surgery+sofa, data = total, family = binomial)
sum_res6 <- summary(res6)
#sum_res6
```

```{r}
invfisher.logit <- vcov(res6)
CI.logit = res6$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),
                                              t(t(sqrt(diag(invfisher.logit)))))
out.hos.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.hos.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
rownames(out.hos.f) = c("White female", "Black male", "Black female",'admission_type','age','bmi','charlson_comorbidity_index','cabg','valve_surgery','sofa')
out.hos.f
```

```{r}
# Get the predicted probabilities
predicted <- predict(res6, newdata = total, type = "response")
lower_ci <- predict(res6, type = "response", se.fit = TRUE)$fit - 1.96 * predict(res6, type = "response", se.fit = TRUE)$se.fit 
upper_ci <- predict(res6, type = "response", se.fit = TRUE)$fit + 1.96 * predict(res6, type = "response", se.fit = TRUE)$se.fit
#results for icu2
icu7_table <- aggregate(cbind(predicted, lower_ci, upper_ci) ~ combined, data = total, FUN = mean)
# Print the aggregated data
p_values <- sum_res6$coefficients[2:4, "Pr(>|z|)"]
p_values <- data.frame(p_values = c(NA,p_values))
rownames(p_values) <- c("White male", "White female", "Black male", "Black female")
icu7_table <- cbind(icu7_table,p_values)
print(icu7_table)
```

## icu days 14

```{r}
total$icu_day14 <- factor(ifelse(total$los_icu_day > 14, 1, 0))
```

```{r}
res7 <- glm(icu_day14 ~ combined + admission_type+age+bmi+charlson_comorbidity_index +
                    cabg+valve_surgery+sofa, data = total, family = binomial)
sum_res7 <- summary(res7)
#sum_res7
```

```{r}
invfisher.logit <- vcov(res7)
CI.logit = res7$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),
                                              t(t(sqrt(diag(invfisher.logit)))))
out.hos.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.hos.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
rownames(out.hos.f) = c("White female", "Black male", "Black female",'admission_type','age','bmi','charlson_comorbidity_index','cabg','valve_surgery','sofa')
out.hos.f
```


```{r}
# Get the predicted probabilities
predicted <- predict(res7, newdata = total, type = "response")
lower_ci <- predict(res7, type = "response", se.fit = TRUE)$fit - 1.96 * predict(res7, type = "response", se.fit = TRUE)$se.fit 
upper_ci <- predict(res7, type = "response", se.fit = TRUE)$fit + 1.96 * predict(res7, type = "response", se.fit = TRUE)$se.fit
#results for icu2
icu14_table <- aggregate(cbind(predicted, lower_ci, upper_ci) ~ combined, data = total, FUN = mean)
# Print the aggregated data
p_values <- sum_res7$coefficients[2:4, "Pr(>|z|)"]
p_values <- data.frame(p_values = c(NA,p_values))
rownames(p_values) <- c("White male", "White female", "Black male", "Black female")
icu14_table <- cbind(icu14_table,p_values)
print(icu14_table)
```



```{r}
icu2_table$icu <- "2"
icu3_table$icu <- "3"
icu7_table$icu <- "7"
icu14_table$icu <- "14"

combined_data2 <- bind_rows(icu2_table,icu3_table,icu7_table,icu14_table)
p_values_filtered <- combined_data2[!is.na(combined_data2$p_values), ]

combined_data2$result_predicted <- (1-combined_data2$predicted)*100
combined_data2$result_lower_ci <- (1-combined_data2$upper_ci)*100
combined_data2$result_upper_ci <- (1-combined_data2$lower_ci)*100

spacing <- 0.06
```

```{r}
line_icu <- ggplot(combined_data2, aes(x = factor(icu, level = c('2', '3', '7', '14')), y = result_predicted, group = combined, color = combined)) +
  geom_errorbar(aes(ymin = result_lower_ci, ymax = result_upper_ci), width = 0.1) +
  geom_line(size = 1) +
  labs(x = "ICU stay duration (day)", y = "ICU transfer rate (%)") +
  scale_color_manual(name = NULL,values = my_palette) +
  scale_y_continuous(limits = c(0, 110), expand = c(0, 0)) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 12),
#    legend.position = "bottom",
#    legend.justification = c(1, 0),
#    legend.box.just = "right",
    legend.position = "none",            # Remove legend
    panel.background = element_blank(),
    panel.grid.major = element_blank(),    # Remove major vertical grid lines
    panel.grid.minor = element_blank(),    # Remove minor vertical grid lines
    panel.border = element_blank(),
    axis.line = element_line(colour = "black", size = 0.5),
    panel.grid.major.y = element_line(colour = "grey", linetype = "solid", size = 0.2)) 
#+ theme(legend.position =  c(0.8, 0.25))
line_icu 
```





2 day here means the probability of each subgroup leaving icu within 2 days, same for 3 days, 7 days, and 14 days.


### hopital days 3

```{r}
res8 <- glm(hos_day3 ~ combined + admission_type+age+bmi+charlson_comorbidity_index +
                    cabg+valve_surgery+sofa, data = total, family = binomial)
sum_res8 <- summary(res8)
#sum_res8
```

```{r}
invfisher.logit <- vcov(res8)
CI.logit = res8$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),
                                              t(t(sqrt(diag(invfisher.logit)))))
out.hos.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.hos.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
rownames(out.hos.f) = c("White female", "Black male", "Black female",'admission_type','age','bmi','charlson_comorbidity_index','cabg','valve_surgery','sofa')
out.hos.f
```


```{r}
# Get the predicted probabilities
predicted <- predict(res8, newdata = total, type = "response")
lower_ci <- predict(res8, type = "response", se.fit = TRUE)$fit - 1.96 * predict(res8, type = "response", se.fit = TRUE)$se.fit 
upper_ci <- predict(res8, type = "response", se.fit = TRUE)$fit + 1.96 * predict(res8, type = "response", se.fit = TRUE)$se.fit
#results for hos3
hos3_table <- aggregate(cbind(predicted, lower_ci, upper_ci) ~ combined, data = total, FUN = mean)
# Print the aggregated data
p_values <- sum_res8$coefficients[2:4, "Pr(>|z|)"]
p_values <- data.frame(p_values = c(NA,p_values))
rownames(p_values) <- c("White male", "White female", "Black male", "Black female")
hos3_table <- cbind(hos3_table,p_values)
print(hos3_table)
```

## hospital days 7

```{r}
total$hos_day7 <- factor(ifelse(total$los_hospital_day > 7, 1, 0))
```

```{r}
res9 <- glm(hos_day7 ~ combined + admission_type+age+bmi+charlson_comorbidity_index +
                    cabg+valve_surgery+sofa, data = total, family = binomial)
sum_res9 <- summary(res9)
#sum_res9
```

```{r}
invfisher.logit <- vcov(res9)
CI.logit = res9$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),
                                              t(t(sqrt(diag(invfisher.logit)))))
out.hos.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.hos.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
rownames(out.hos.f) = c("White female", "Black male", "Black female",'admission_type','age','bmi','charlson_comorbidity_index','cabg','valve_surgery','sofa')
out.hos.f
```

```{r}
# Get the predicted probabilities
predicted <- predict(res9, newdata = total, type = "response")
lower_ci <- predict(res9, type = "response", se.fit = TRUE)$fit - 1.96 * predict(res9, type = "response", se.fit = TRUE)$se.fit 
upper_ci <- predict(res9, type = "response", se.fit = TRUE)$fit + 1.96 * predict(res9, type = "response", se.fit = TRUE)$se.fit
#results for hos3
hos7_table <- aggregate(cbind(predicted, lower_ci, upper_ci) ~ combined, data = total, FUN = mean)
# Print the aggregated data
p_values <- sum_res9$coefficients[2:4, "Pr(>|z|)"]
p_values <- data.frame(p_values = c(NA,p_values))
rownames(p_values) <- c("White male", "White female", "Black male", "Black female")
hos7_table <- cbind(hos7_table,p_values)
print(hos7_table)
```



## hospital days 14

```{r}
total$hos_day14 <- factor(ifelse(total$los_hospital_day > 14, 1, 0))
```

```{r}
res10 <- glm(hos_day14 ~ combined + admission_type+age+bmi+charlson_comorbidity_index +
                    cabg+valve_surgery+sofa, data = total, family = binomial)
sum_res10 <- summary(res10)
#sum_res10
```

```{r}
invfisher.logit <- vcov(res10)
CI.logit = res10$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),
                                              t(t(sqrt(diag(invfisher.logit)))))
out.hos.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.hos.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
rownames(out.hos.f) = c("White female", "Black male", "Black female",'admission_type','age','bmi','charlson_comorbidity_index','cabg','valve_surgery','sofa')
out.hos.f
```

```{r}
# Get the predicted probabilities
predicted <- predict(res10, newdata = total, type = "response")
lower_ci <- predict(res10, type = "response", se.fit = TRUE)$fit - 1.96 * predict(res10, type = "response", se.fit = TRUE)$se.fit 
upper_ci <- predict(res10, type = "response", se.fit = TRUE)$fit + 1.96 * predict(res10, type = "response", se.fit = TRUE)$se.fit
#results for hos3
hos14_table <- aggregate(cbind(predicted, lower_ci, upper_ci) ~ combined, data = total, FUN = mean)
# Print the aggregated data
p_values <- sum_res10$coefficients[2:4, "Pr(>|z|)"]
p_values <- data.frame(p_values = c(NA,p_values))
rownames(p_values) <- c("White male", "White female", "Black male", "Black female")
hos14_table <- cbind(hos14_table,p_values)
print(hos14_table)
```

## hospital days 30

```{r}
total$hos_day30 <- factor(ifelse(total$los_hospital_day > 30, 1, 0))
```

```{r}
res11 <- glm(hos_day30 ~ combined + admission_type+age+bmi+charlson_comorbidity_index +
                    cabg+valve_surgery+sofa, data = total, family = binomial)
sum_res11 <- summary(res11)
#sum_res11
```

```{r}
invfisher.logit <- vcov(res11)
CI.logit = res11$coefficients + kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))),
                                              t(t(sqrt(diag(invfisher.logit)))))
out.hos.f = cbind(exp(CI.logit)[-1,,drop=FALSE])
colnames(out.hos.f) = c('Estimate for OR','95% CI Lower','95% CI Upper')
rownames(out.hos.f) = c("White female", "Black male", "Black female",'admission_type','age','bmi','charlson_comorbidity_index','cabg','valve_surgery','sofa')
out.hos.f
```

```{r}
# Get the predicted probabilities
predicted <- predict(res11, newdata = total, type = "response")
lower_ci <- predict(res11, type = "response", se.fit = TRUE)$fit - 1.96 * predict(res11, type = "response", se.fit = TRUE)$se.fit 
upper_ci <- predict(res11, type = "response", se.fit = TRUE)$fit + 1.96 * predict(res11, type = "response", se.fit = TRUE)$se.fit
#results for hos3
hos30_table <- aggregate(cbind(predicted, lower_ci, upper_ci) ~ combined, data = total, FUN = mean)
# Print the aggregated data
p_values <- sum_res11$coefficients[2:4, "Pr(>|z|)"]
p_values <- data.frame(p_values = c(NA,p_values))
rownames(p_values) <- c("White male", "White female", "Black male", "Black female")
hos30_table <- cbind(hos30_table,p_values)
print(hos30_table)
```


```{r}
hos3_table$hos <- "3"
hos7_table$hos <- "7"
hos14_table$hos <- "14"
hos30_table$hos <- "30"

combined_data3 <- bind_rows(hos3_table,hos7_table,hos14_table,hos30_table)
p_values_filtered <- combined_data3[!is.na(combined_data3$p_values), ]


combined_data3$result_predicted <- (1-combined_data3$predicted)*100
combined_data3$result_lower_ci <- (1-combined_data3$upper_ci)*100
combined_data3$result_upper_ci <- (1-combined_data3$lower_ci)*100


spacing <- 0.04
```

```{r}
line_hospital <- ggplot(combined_data3, aes(x = factor(hos, level = c('3', '7', '14', '30')), y = result_predicted, group = combined, color = combined)) +
  geom_errorbar(aes(ymin = result_lower_ci, ymax = result_upper_ci), width = 0.1) +
  geom_line(size = 1) +
  labs(x = "Hospital stay duration (day)", y = "Hospital discharge rate (%)") +
  scale_color_manual(name = NULL, values = my_palette) +
  scale_y_continuous(limits = c(0, 110), expand = c(0, 0)) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 12),
#    legend.position = "bottom",          # Set legend position to bottom
 #   legend.justification = c(1, 1),      # Set legend justification to bottom right
#    legend.box.just = "right",
    legend.position = "none",            # Remove legend
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.line = element_line(colour = "black", size = 0.5),
    panel.grid.major.y = element_line(colour = "grey", linetype = "solid", size = 0.2)) 
#+ theme(legend.position =  c(0.8, 0.25))
line_hospital 
```



3 day here means the probability of each subgroup leaving hospital within 3 days, same for 7 days, 14 days, and 30 days.

```{r}
#install.packages("ggpubr")
library(ggpubr)
line_combined <- ggarrange(line_mv, line_icu, line_hospital,
                  #labels = c("a)", "b)", "c)"),
                     common.legend = FALSE, ncol = 3, nrow = 1)
ggsave("line_combined.pdf", line_combined,width = 10, height = 3)
```


#install.packages("cowplot")
library(cowplot)

line_combined <- grid.arrange(line_mv, line_icu, line_hospital, nrow = 1,
             widths = c(0.2, 0.2, 0.2),  # Adjust the width proportions
             heights = c(0.1))

ggsave("line_combined.pdf", line_combined,width = 10, height = 3)




## Conclusion
Corresponding to PSM study, black female, black male and whilte female tend to have longer mv hours, icu days, and hospital days compare to black male with adjusted confounders: age, admission type, charlson comorbidity index, cabg, valve surgery and sofa score. 

There are no significant difference for black male, white male for icu days of 7 and 14, no significant difference between all four subgroups for hospital day 3 and no significant difference between black female, black male, and white male for hospital days of 30, and no significant difference between black male and white male for ventilation hours of 0 meaning they stopped ventilation as soon as they get into icu.
