---
title: "inequality_5"
author: "Zhongyun Zhang"
date: "2023-08-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tableone)
library(tidyverse)
library(MatchIt)
library(lmtest)
library(sandwich)
library(dplyr)
library("viridis")
#install.packages("gtools", repos="https://cran.rstudio.com/")
library(gtools)
```

## Data Processing

* Outcomes: vent hours, hospital days, icu days
* Controlling variables: age, CCI score, SOFA score, Surgery type (cabg, valve)
* Varaibles: race (black and white), gender

```{r}
#import data
eicu <- read.csv("~/Desktop/301/PPC/data/hf_eicu.csv",row.names = 1, header = TRUE)
mimic <- read.csv("~/Desktop/301/PPC/data/hf_mimic.csv", row.names = 1, header = TRUE)
#View(eicu)
```

```{r}
#selecting eicu number of beds>250 
eicu <- subset(eicu,eicu$numbedscategory==">= 500" | eicu$numbedscategory=="250 - 499")
table(eicu$ethnicity)
table(mimic$ethnicity)
```

```{r}
## fill na for MV hours
eicu[is.na(eicu)] <- 0
mimic[is.na(mimic)] <- 0
# Remove rows with one patient with 300 in 'age' - mimic
mimic <- mimic[mimic$age != "300", ]
```

```{r}
#select variables
vars <- c('admission_type', 'gender', 'age',"bmi", 'charlson_comorbidity_index', 'ethnicity', 'cabg', 'valve_surgery', 'sofa', 'vent_hour', 'los_hospital_day', 'los_icu_day')
eicu <- eicu %>% select(all_of(vars))
mimic <- mimic %>% select(all_of(vars))
# combine eicu and mimic
total <- rbind(eicu, mimic) %>%
  filter(ethnicity %in% c('black', 'white'))
#hist(total$ethnicity,total$age)
total <- total[total$vent_hour >= 0,]
#table(total$ethnicity)
```


```{r}
# Assuming your dataset is named 'data' and 'gender' and 'ethnicity' are existing columns
total$combined <- ifelse(total$gender == "M" & total$ethnicity == "white", "White male",
                 ifelse(total$gender == "F" & total$ethnicity == "white", "White female",
                 ifelse(total$gender == "M" & total$ethnicity == "black", "Black male",
                 ifelse(total$gender == "F" & total$ethnicity == "black", "Black female",NA))))
# Convert 'combined' column to numeric representation
total$combined <- factor(total$combined, levels = c("White male", "White female", "Black male", "Black female"), labels = c(1, 2, 3, 4))
levels(total$combined) <- c("White male", "White female", "Black male", "Black female")
table(total$combined)
```


total$age_group <-  ifelse(total$age > 18 & total$age <= 45, "18-45",
                 ifelse(total$age > 45 & total$age <= 65, "45-65",
                 ifelse(total$age > 65 & total$age <= 75, "65-75",">75")))
```{r}
total$bmi_c <-  ifelse(total$bmi < 18.5, "Underweight",
                 ifelse(total$bmi >= 18.5 & total$bmi < 25, "Healthy weight",
                 ifelse(total$bmi >= 25 & total$bmi < 30, "Overweight",
                 ifelse(total$bmi >= 30 & total$bmi < 40, "Obese",                                             "Extremely obese"))))


total$mv_group <-  ifelse(total$vent_hour == 0.000, "0",
                 ifelse(total$vent_hour > 0 & total$vent_hour <= 12, "12",
                 ifelse(total$vent_hour > 12 & total$vent_hour <= 24, "24",
                 ifelse(total$vent_hour > 24 & total$vent_hour <= 48,"48",
                 ifelse(total$vent_hour > 48 & total$vent_hour <= 168, "168",">168")))))

total$los_icu_group <-  ifelse(total$los_icu_day <= 2, "2",
                 ifelse(total$los_icu_day > 2 & total$los_icu_day <= 3, "3",
                 ifelse(total$los_icu_day > 3 & total$los_icu_day <= 7, "7",
                 ifelse(total$los_icu_day > 7 & total$los_icu_day <= 14, "14",">14"))))

total$los_hospital_group <-  ifelse(total$los_hospital_day <= 3, "3",
                 ifelse(total$los_hospital_day > 3 & total$los_hospital_day <= 7, "7",
                 ifelse(total$los_hospital_day > 7 & total$los_hospital_day <= 14, "14",
                 ifelse(total$los_hospital_day > 14 & total$los_hospital_day <= 30, "30", 
                        ">30"))))
```

```{r}
# Reorder levels of age_group variable
#total$age_group <- factor(total$age_group, levels = c("18-45", "45-65", "65-75",">75"))

# Reorder levels of mv_group variable
total$mv_group <- factor(total$mv_group, levels = c("0", "12", "24", "48","168",">168"))

# Reorder levels of los_icu_group variable
total$los_icu_group <- factor(total$los_icu_group, levels = c("2", "3", "7", "14", ">14"))

# Reorder levels of los_hospital_group variable
total$los_hospital_group <- factor(total$los_hospital_group, levels = c("3", "7", "14", "30", ">30"))

total$bmi_c <- factor(total$bmi_c, levels = c("Underweight", "Healthy weight", "Overweight", "Obese", "Extremely obese"))
```

## Table One

```{r}
table1 <- CreateTableOne(vars = c('age',"bmi_c",
                                  'admission_type','charlson_comorbidity_index', 'sofa',
                                  'cabg','valve_surgery', 'mv_group',
                                  'los_icu_group', 'los_hospital_group'),
                         data = total,
                         factorVars = c('admission_type', 'cabg',"bmi_c",
                                        'valve_surgery', 'mv_group', 'los_icu_group',
                                        'los_hospital_group'),
                         strata = 'combined',
                         smd = TRUE)

table1 <- print(table1, smd=TRUE, showAllLevels = TRUE, noSpaces = TRUE, 
                printToggle = FALSE, nonnormal = c("charlson_comorbidity_index", 'sofa','age'))

write.csv(table1, file = "~/Desktop/301/PPC/output/table1_4subgroup.csv")
```


#total
```{r}
table1 <- CreateTableOne(vars = c('age',"bmi_c",
                                  'admission_type','charlson_comorbidity_index', 'sofa',
                                  'cabg','valve_surgery', 'mv_group',
                                  'los_icu_group', 'los_hospital_group'),
                         data = total,
                         factorVars = c('admission_type', 'cabg',"bmi_c",
                                        'valve_surgery', 'mv_group', 'los_icu_group',
                                        'los_hospital_group'),
                         smd = TRUE)

table1 <- print(table1, smd=TRUE, showAllLevels = TRUE, noSpaces = TRUE, 
                printToggle = FALSE, nonnormal = c("charlson_comorbidity_index", 'sofa','age'))

write.csv(table1, file = "~/Desktop/301/PPC/output/table1_total.csv")
```


## ZHIHUIJINGHUA

### MV

add_stars <- function(p_values, threshold = 0.05) {
  if (p_values < threshold) {
    return(paste0(format(p_values, scientific = FALSE), "***"))
  } else {
    return(format(p_values, scientific = FALSE))
  }
}

# Add stars to p-values less than threshold using mutate()
p_values_filtered <- p_values_filtered %>% mutate(starred_p_value = sapply(p_values, add_stars, threshold = 0.05))

ggplot(combined_data, aes(x = factor(mv, level=c('0 hour', '12 hour', '24 hour', '48 hour')) , y = (1-predicted), group = combined, color = combined)) +
  geom_errorbar(aes(ymin = (1-upper_ci), ymax = (1-lower_ci)), width = 0.1) +
  geom_line(size = 0.8) +
  labs(x = "MV Hour", y = "MV Removal Rate") +
  scale_color_discrete(name = "Combined") + 
  scale_color_manual(values = my_palette) +
  theme_minimal() +
  guides(color = guide_legend(title = "")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
    legend.position = "right" ) +
  geom_text(data = p_values_filtered, aes(x = factor(mv, level=c('0 hour', '12 hour', '24 hour', '48 hour')), y = max(p_values_filtered$predicted)+spacing *(1:nrow(p_values_filtered)), group = mv,label = paste(p_values_filtered$starred_p_value)), size = 3) +
 coord_cartesian(ylim = c(0, 1.5), expand = TRUE)  # Adjust the y-axis limits to include p-values
 
### ICU day


p_values_filtered <- p_values_filtered %>% 
  mutate(p_values = round(p_values, digits = 8))
# Function to add stars to p-values less than a threshold
add_stars <- function(p_values, threshold = 0.05) {
  if (p_values < threshold) {
    return(paste0(format(p_values, scientific = FALSE), "***"))
  } else {
    return(format(p_values, scientific = FALSE))
  }
}
# Add stars to p-values less than threshold using mutate()
p_values_filtered <- p_values_filtered %>% mutate(starred_p_value = sapply(p_values, add_stars, threshold = 0.05))


ggplot(combined_data2, aes(x = factor(icu, level=c('2 day', '3 day', '7 day', '14 day')) , y = (1-predicted), group = combined, color = combined)) +
  geom_errorbar(aes(ymin = (1-upper_ci), ymax = (1-lower_ci)), width = 0.1) +
  geom_line(size = 0.8) +
  labs(x = "ICU Days", y = "ICU Leaving Rate") +
  scale_color_discrete(name = "Combined") + 
  scale_color_manual(values = my_palette) +
  theme_minimal() +
  guides(color = guide_legend(title = "")) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
    legend.position = "right") +
  geom_text(data = p_values_filtered, aes(x = factor(icu,, level=c('2 day', '3 day', '7 day', '14 day')), y = max(p_values_filtered$predicted)+spacing *(1:nrow(p_values_filtered)),label = paste(p_values_filtered$starred_p_value)), size = 3) +
 coord_cartesian(ylim = c(0, 1.5), expand = TRUE)  
 
 
## Hospital Days

p_values_filtered <- p_values_filtered %>% 
  mutate(p_values = round(p_values, digits = 8))
# Function to add stars to p-values less than a threshold
add_stars <- function(p_values, threshold = 0.05) {
  if (p_values < threshold) {
    return(paste0(format(p_values, scientific = FALSE), "***"))
  } else {
    return(format(p_values, scientific = FALSE))
  }
}
combined_data3 <- bind_rows(hos3_table,hos7_table,hos14_table,hos30_table)
p_values_filtered <- combined_data3[!is.na(combined_data3$p_values), ]
# Add stars to p-values less than threshold using mutate()
p_values_filtered <- p_values_filtered %>% mutate(starred_p_value = sapply(p_values, add_stars, threshold = 0.05))


ggplot(combined_data3, aes(x = factor(hos, level=c('3 day', '7 day', '14 day', '30 day')) , y = (1-predicted), group = combined, color = combined)) +
  geom_errorbar(aes(ymin = (1-upper_ci), ymax = (1-lower_ci)), width = 0.1) +
  geom_line(size = 0.8) +
  labs(x = "Hospital Days", y = "Hospital Leaving Rate") +
  scale_color_discrete(name = "Combined") + 
  scale_color_manual(values = my_palette) +
  theme_minimal() +
  guides(color = guide_legend(title = "")) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
    legend.position = "right") +
  geom_text(data = p_values_filtered, aes(x = factor(hos, level=c('3 day', '7 day', '14 day', '30 day')), y = max(p_values_filtered$predicted)+spacing *(1:nrow(p_values_filtered)),label = paste(p_values_filtered$starred_p_value)), size = 3) +
 coord_cartesian(ylim = c(0, 1.5), expand = TRUE)  

